Forecasting short-term hourly Emergency Departement arrivals

### 2021-10-28 paper/paper.Rmd

plot_data <- data.frame(Hour = 0:23)
for(i in c("Mon","Tue","Wed","Thu","Fri","Sat","Sun","Public Holiday","Christmas Day","New Years Day")){
  pef <- gamlss::getPEF(h2_gamlss2,term="clock_hour",parameter = "mu",fixed.at = list(dow3=i))
  plot_data[[i]] <- pef(0:23)
}

plot_data <- reshape::melt.data.frame(plot_data,id.vars="Hour",variable_name="Day Type")

plot_1 <- ggplot(plot_data,aes(x=Hour,y=value,color=`Day Type`)) + geom_smooth(se = F,span=0.5) +
  ylab("Effect [Expected Arrivals]") + theme_few()

# plot_1

plot_data2 <- data.frame(Hour = 0:23)
for(i in c("Mon","Tue","Wed","Thu","Fri","Sat","Sun","Public Holiday","Christmas Day","New Years Day")){
  pef <- gamlss::getPEF(h2_gamlss2,term="clock_hour",parameter = "sigma",fixed.at = list(dow3=i))
  plot_data2[[i]] <- pef(0:23)
}

plot_data2 <- reshape::melt.data.frame(plot_data2,id.vars="Hour",variable_name="Day Type")

plot_2 <- ggplot(plot_data2,aes(x=Hour,y=value,color=`Day Type`)) + geom_smooth(se = F,span=0.5) +
  ylab("Effect [Standard Deviation of Arrivals]") + theme_few()

#plot_2

grid.arrange(plot_1, plot_2, ncol=2)

## Forecast evaluation

Probabilistic forecasts are evaluated following the principle of *sharpness subject to calibration*, meaning that the sharper forecast is prefered provided that it is calibrated. Mis-calibrated forecasts are unsuitable for use in decision-making so should be excluded. Calibration is evaluated visually in \@ref(fig-quantilebias), which highlights a systematic negative bias accross all probability levels in many models, with only the truncated normal and $t$ family GAMLSS models (NOtr-1, NOtr-2, Ttr-2) and iETS models showing good calibration across most probability levels. Notably, both benchmarks exhibit negative quantile bias as they struggle to capture the long term trend of increasing attendance.

```{r fig-quantilebias, fig.cap= "Quantile ...", fig.align='center',out.width= "90%"}
#knitr::include_graphics(rep("../results/QuantileBias.png"))
## Should reduce the number of models shown here... it's really hard to tell them appart!

REL_nom <- data.table(Nominal=seq(0,1,by=0.05),
                      Empirical=seq(0,1,by=0.05),
                      `Quantile Bias`= 0,
                      Method="Nominal")

REL[,`Quantile Bias`:=Empirical-Nominal]

ggplot(data=REL[Horizon=="All" & Issue == "All",],aes(x=Nominal,y=`Quantile Bias`,group=Method,color=Method)) +
  geom_line(data=REL_nom,aes(x=Nominal,y=`Quantile Bias`), color="black",size=1.1,show.legend = F) +
  geom_line() + geom_point() +
  xlim(c(0,1)) + ylim(c(-0.2,0.2)) + #ggtitle("Quantile Bias") +
  theme_few() 
```

```{r tab-results, out.width= "50%"}
results_table <- read_rds("results_table.rds")
knitr::kable(results_table, booktabs = T, 
             linesep = "",caption = "Summary of studies in hourly emergency care forecasting") %>% kableExtra::kable_styling(latex_options = c("hold_position"), font_size =11)
```

Evaluation metrics from the test period are presented in \@ref(tab-results) which are ordered by Quantile Bias. The five models identified above have a Quantile Bias of 0.014 or less, which is substantially lower than the next group of forecast with Quantile Biases of 0.037 and above, ETS(XXX) being the only exception with a value of 0.019.

```{r tab-results-qbias, out.width= "50%"}
results_table_qb <- read_rds("results_table.rds") %>% select(`Quantile Bias`)

knitr::kable(results_table_qb, booktabs = T, 
             linesep = "",caption = "Forecast evluation for all methods on test data in decreasing order by Quantile Bias") %>% kableExtra::kable_styling(latex_options = c("hold_position"), font_size =11)
```

The Pinball Score evaluates overall performance, penalising both poor calibration and lack of sharpness, and ranks the various models quite differently. NBI-2 and Poisson-2 achive the lowest Pinball Loss, but as we have seen, are poorly calibrated so not suitable for use in decision making. Of the well calibrated methods, NOtr-2 and Ttr-2 have the lowest Pinball Score, and provide a significant improvement over Benchmark 2 by this metric. However, we note that while the improvement in terms of Pinball Score appears small, these methods have substantially improved forecast calibration compared to the benchmarks, which are both poorly calibrated.

```{r fig-skill, fig.cap= "Skill score ...", fig.align='center',out.width= "90%"}
knitr::include_graphics(rep("../results/Skill_rel2bench_reduced.png"))

# knitr::include_graphics(rep("../results/Skill_rel2bench_reduced.png"))

load("../results/all_results_paper_2021-10-07.Rda")

plotdata <- melt(bootdata[,100*(get(REF)-.SD)/get(REF),.SDcols=NAMES],
                 measure.vars =  1:length(NAMES),variable.name = "Method",
                 value.name = "Skill Score")

## Merge Reliability for colouring...
plotdata <- merge(plotdata,REL[kfold=="Test" & Horizon=="All" & Issue == "All",.(Qbias=mean(abs(Nominal-Empirical))),by="Method"],
                  by="Method",all.x = T)

ggplot(plotdata[`Skill Score`>-2,], aes(x=reorder(Method, -`Skill Score`), y=`Skill Score`, fill=Qbias)) + 
  ylab("Pinball Skill Score [%]") +
  geom_boxplot() + theme_few() +
  # ggtitle("Pinball Skill Score Relative to Benchmark 2") +
  theme(axis.text.x = element_text(angle = -80)) + #scale_fill_manual(values=cbPalette) +
  # scale_x_discrete(labels= paste0(substring(NAMES, 1,4),".")) +
  geom_hline(yintercept=0, linetype="dashed",size=0.5)+
  scale_fill_gradientn(colours = rev(brewer.pal(n = 11, name = "RdYlGn")),
                       limits=c(0,0.11))+
  labs(fill = "Quantile Bias",x="Method")

Forecast performance by time of day...

@@

>>>>

Evaluation metrics from all Methods on the test period are presented in \@ref(tab-results) which are ordered by Quantile Bias. The five models identified above have a Quantile Bias of 0.014 or less, which is substantially lower than the next group of forecast with Quantile Biases of 0.037 and above.

```{r tab-results, out.width= "50%"}
results_table <- read_rds("results_table.rds")
results_table <- results_table[!results_table$Method%in%c("ADAM-iETSX-Ceiling","LinearRegression","Poisson-2-I","fasster","qreg-1"),]
results_table <- results_table[order(results_table$`Quantile Bias`),]

knitr::kable(results_table, booktabs = T, 
             linesep = "",caption = "Summary of studies in hourly emergency care forecasting") %>% kableExtra::kable_styling(latex_options = c("hold_position"), font_size =11)
```

The Pinball Score evaluates overall performance, penalising both poor calibration and lack of sharpness, and ranks the various models quite differently. Pinball is visualised in Figure \'ref(Pinball) broken down by quantile. NBI-2 and Poisson-2 achieve the lowest Pinball Loss, but are poorly calibrated so not suitable for use in decision making. Of the well calibrated methods, NOtr-2 and Ttr-2 have the lowest Pinball Score, and provide a significant improvement over Benchmark 2 by this metric. However, we note that while the improvement in terms of Pinball Score appears small, these methods have substantially improved forecast calibration compared to the benchmarks, which are poorly calibrated.

```{r Pinball, fig.cap= "Pinball loss for selected forecsating methods and quantiles from 5% to 95%. Several methods perform similarly well relative to others by this metric but calibration should also be considered when evaluating probabilistic foreasts.", fig.align='center',fig.width= 7,fig.height=6}
```

In order to verify that this improvement is a result of improved modelling and not a result of sample variation, we have performed block bootstrap resampling on Pinball skill scores relative to Bechmark-2 for each forecast. The results for all methods with a skill score greater than -2% are shown in Figure \@ref(Skill_rel2bench_reduced), where boxes are coloured according to calibration. Where boxes do not overlap we can be confident that apparent differences in skill are not the result of sample variation.

```{r Skill_rel2bench_reduced, fig.cap= "Bootstrapped skill scores for top-performing methods relative to Benchmark-2. Non-overlapping boxes indicate that we can be confident that apparent differences in Pinball Score are not the result of sample variation.", fig.align='center',fig.width= 7,fig.height=6}
```

```{r Pinaball_vs_Qbias, fig.cap= "Pinball Score vs average Qunatile Bias for selected methods. While some methods perform well by one measure or the other, only NOtr-2 is both calibrated and has a low Pinball Score.", fig.align='center',fig.width= 7,fig.height=6}
```

We delve deeper into forecast performance by evaluating forecasts by time of day, and lead-time. The Pinball Score for slected methods is plotted against lead-time in Figure \@ref(lead-time-pb), separated by whether the forecast was issued at midnight (00) or mid-day (12). We can see clearly that time of day is a greater influence on performance than lead-time, with morning arrivals and the evening peak in arrivals proving the most difficult to predict.

```{r lead-time-pb, fig.cap= "Pinball Score by lead-time for forecasts issued at midnight (00) and mid-day (12).", fig.align='center',fig.width= 7,fig.height=6}
```

Computational time should also be considered, as national public health bodies may be required to produce forecasts of many variables for a large number of hospitals on  regular basis...

@@

1. Conditional distribution of future arrivals is neither Gaussian nor Poisson distributed, and requires careful treatment.

2. Simple benchmarks and "out-of-the-box" forecasting methods produce poor predictions of A&E arrivals.

3. A distributional regression appraoch, with additive models for both the location and scale parameters of a truncated Normal distirbution produces best all-round forecast performance
